---
output:
  html_document: default
  word_document: default
---

```{r}
ecommerce <- read.csv('online_shoppers_intention.csv')
#3. Checking the data

```
```{r}
#Viewing the first 6 entries
head(ecommerce)
```
```{r}
getwd()
```
```{r}
#viewing the last 6 entries
tail(ecommerce)
```
```{r}
#viewing the structure of the dataset
str(ecommerce)
```
```{r}
#Finding the number of columns and rows 
dim(ecommerce)
```
```{r}
#Statistical summary of our data set
summary(ecommerce)
```
#4. Tidying the data/ Data Cleaning

## Completeness
```{r}
#checking for missing values
colSums(is.na(ecommerce))
```

```{r}
#Removing the missing values
ec <- na.omit(ecommerce)
colSums(is.na(ec))
dim(ec)
```
## consistency
```{r}
#checking for duplicates
anyDuplicated(ec)
```
```{r}
#removing duplicates
ec <- unique(ec)
dim(ec)
```
```{r}
#convert categorical columns to factors
ec[,11:18] <- sapply(ec[,11:18], as.factor)
head(ec)
```

```{r}
#checking for outliers
#list(colnames(ec))
boxplot(ec$Administrative, main= 'Boxplot of administrative web pages', xlab='administrative', ylab='value')
boxplot(ec$Administrative_Duration, main= 'Boxplot of administrative web pages duration', xlab='administrative duration', ylab='value')
boxplot(ec$Informational, main= 'Boxplot of informational web pages', xlab='informational', ylab='value')
boxplot(ec$Informational_Duration, main= 'Boxplot of informational web pages duration', xlab='informational duration', ylab='value')
boxplot(ec$ProductRelated, main= 'Boxplot of product related web pages', xlab='product related', ylab='value')
boxplot(ec$ProductRelated_Duration, main= 'Boxplot of product related web pages duration', xlab='product related duration', ylab='value')
boxplot(ec$BounceRates, main= 'Boxplot of bounce rates', xlab='bounce rates', ylab='value')
boxplot(ec$ExitRates, main= 'Boxplot of exit rates', xlab='exit rates', ylab='value')
boxplot(ec$PageValues, main= 'Boxplot of page values', xlab='page values', ylab='value')
boxplot(ec$SpecialDay, main= 'Boxplot of special day', xlab='special day', ylab='value')
```
```There are outliers in every column but we wunt remove them as they might give insight to the project.```
#5. Univariate Exploratory Data Analysis
## Measures of central tendancy
```{r}
#finding the mean
mean <- colMeans(ec[sapply(ec, is.numeric)])
mean
#Finding the median
#loading the tidyverse and robustbase(for the colMedians function) libraries
library(robustbase)
library(tidyverse)
median <- ec%>%
  select_if(is.numeric) %>%
  as.matrix()%>%
  colMedians()
print(median)
#Finding the mode
mode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
mode(ec$Administrative)
mode(ec$Administrative_Duration)
mode(ec$Informational)
mode(ec$Informational_Duration)
mode(ec$ProductRelated)
mode(ec$ProductRelated_Duration)
mode(ec$BounceRates)
mode(ec$ExitRates)
mode(ec$PageValues)
mode(ec$SpecialDay)
```
##Measures of dispersion
```{r}
#finding the minimum
min(ec$Administrative)
min(ec$Administrative_Duration)
min(ec$Informational)
min(ec$Informational_Duration)
min(ec$ProductRelated)
min(ec$ProductRelated_Duration)
min(ec$BounceRates)
min(ec$ExitRates)
min(ec$PageValues)
min(ec$SpecialDay)
#finding the maximum
max(ec$Administrative)
max(ec$Administrative_Duration)
max(ec$Informational)
max(ec$Informational_Duration)
max(ec$ProductRelated)
max(ec$ProductRelated_Duration)
max(ec$BounceRates)
max(ec$ExitRates)
max(ec$PageValues)
max(ec$SpecialDay)
#finding the range
range(ec$Administrative)
range(ec$Administrative_Duration)
range(ec$Informational)
range(ec$Informational_Duration)
range(ec$ProductRelated)
range(ec$ProductRelated_Duration)
range(ec$BounceRates)
range(ec$ExitRates)
range(ec$PageValues)
range(ec$SpecialDay)
#finding the quantile
quantile(ec$Administrative)
quantile(ec$Administrative_Duration)
quantile(ec$Informational)
quantile(ec$Informational_Duration)
quantile(ec$ProductRelated)
quantile(ec$ProductRelated_Duration)
quantile(ec$BounceRates)
quantile(ec$ExitRates)
quantile(ec$PageValues)
quantile(ec$SpecialDay)
#finding the standard deviation
sd(ec$Administrative)
sd(ec$Administrative_Duration)
sd(ec$Informational)
sd(ec$Informational_Duration)
sd(ec$ProductRelated)
sd(ec$ProductRelated_Duration)
sd(ec$BounceRates)
sd(ec$ExitRates)
sd(ec$PageValues)
sd(ec$SpecialDay)
#finding variance
var(ec$Administrative)
var(ec$Administrative_Duration)
var(ec$Informational)
var(ec$Informational_Duration)
var(ec$ProductRelated)
var(ec$ProductRelated_Duration)
var(ec$BounceRates)
var(ec$ExitRates)
var(ec$PageValues)
var(ec$SpecialDay)
#finding kurtosis
library(e1071)
skewness(ec$Administrative)
skewness(ec$Administrative_Duration)
skewness(ec$Informational)
skewness(ec$Informational_Duration)
skewness(ec$ProductRelated)
skewness(ec$ProductRelated_Duration)
skewness(ec$BounceRates)
skewness(ec$ExitRates)
skewness(ec$PageValues)
skewness(ec$SpecialDay)
#finding skewness
kurtosis(ec$Administrative)
kurtosis(ec$Administrative_Duration)
kurtosis(ec$Informational)
kurtosis(ec$Informational_Duration)
kurtosis(ec$ProductRelated)
kurtosis(ec$ProductRelated_Duration)
kurtosis(ec$BounceRates)
kurtosis(ec$ExitRates)
kurtosis(ec$PageValues)
kurtosis(ec$SpecialDay)
```
## Univariate Analysis Graphicals

### Categorical columns
```{r}
#colnames(ec)
#frequency table of month
month.freq <- table(ec$Month)
sort(month.freq, decreasing = TRUE)[1:5]
#Bar chart to show frequency distribution of months 
options(repr.plot.width = 10, repr.plot.height = 10)
barplot(c(month.freq), main="A barchart of months.",
        xlab="months",
        ylab="frequency",
        cex.main=2, cex.lab=1.7,cex.sub=1.2,
        col=c("blue"))
#frequency table of Operating systems
os.freq <- table(ec$OperatingSystems)
sort(os.freq, decreasing = TRUE)[1:5]
#Bar chart to show frequency distribution of operating systems
options(repr.plot.width = 10, repr.plot.height = 10)
barplot(c(os.freq), main="A barchart of operating systems.",
        xlab="operating systems",
        ylab="frequency",
        cex.main=2, cex.lab=1.7,cex.sub=1.2,
        col=c("green"))
#frequency table of browser
browser.freq <- table(ec$Browser)
sort(browser.freq, decreasing = TRUE)[1:5]
#Bar chart to show frequency distribution of browsers 
options(repr.plot.width = 10, repr.plot.height = 10)
barplot(c(browser.freq), main="A barchart of browser.",
        xlab="browser",
        ylab="frequency",
        cex.main=2, cex.lab=1.7,cex.sub=1.2,
        col=c("blue"))
#frequency table of region
region.freq <- table(ec$Region)
sort(region.freq, decreasing = TRUE)[1:5]
#Bar chart to show frequency distribution of regions 
options(repr.plot.width = 10, repr.plot.height = 10)
barplot(c(region.freq), main="A barchart of regions.",
        xlab="region",
        ylab="frequency",
        cex.main=2, cex.lab=1.7,cex.sub=1.2,
        col=c("green"))
#frequency table of traffic type
traffic.freq <- table(ec$TrafficType)
sort(traffic.freq, decreasing = TRUE)[1:5]
#Bar chart to show frequency distribution of traffic type 
options(repr.plot.width = 10, repr.plot.height = 10)
barplot(c(traffic.freq), main="A barchart of traffic type.",
        xlab="traffic type",
        ylab="frequency",
        cex.main=2, cex.lab=1.7,cex.sub=1.2,
        col=c("blue"))
#frequency table of visitor type
visitor.freq <- table(ec$VisitorType)
sort(visitor.freq, decreasing = TRUE)[1:5]
#Bar chart to show frequency distribution of visitor type 
options(repr.plot.width = 10, repr.plot.height = 10)
barplot(c(visitor.freq), main="A barchart of visitor types.",
        xlab="visitor types",
        ylab="frequency",
        cex.main=2, cex.lab=1.7,cex.sub=1.2,
        col=c("green"))
#frequency table of weekend
weekend.freq <- table(ec$Weekend)
sort(weekend.freq, decreasing = TRUE)[1:5]
#Bar chart to show frequency distribution of weekend 
options(repr.plot.width = 10, repr.plot.height = 10)
barplot(c(weekend.freq), main="A barchart of weekend.",
        xlab="weekend",
        ylab="frequency",
        cex.main=2, cex.lab=1.7,cex.sub=1.2,
        col=c("blue", "green"))
#frequency table of revenue
rev.freq <- table(ec$Revenue)
sort(rev.freq, decreasing = TRUE)[1:5]
#Bar chart to show frequency distribution of revenue 
options(repr.plot.width = 10, repr.plot.height = 10)
barplot(c(rev.freq), main="A barchart of revenue.",
        xlab="revenue",
        ylab="frequency",
        cex.main=2, cex.lab=1.7,cex.sub=1.2,
        col=c("blue", "green"))
```
### Numerical variables
```{r}
#colnames(ec)
#creating a histogram of administrative variable.
hist(ec$Administrative)
#creating a histogram of administrative duration variable
hist(ec$Administrative_Duration)
#creating a histogram of informational variable
hist(ec$Informational)
#creating a histogram of informational duration variable
hist(ec$Informational_Duration)
#creating a histogram of product related variable
hist(ec$ProductRelated)
#creating a histogram of product related duration variable
hist(ec$ProductRelated_Duration)
#creating a histogram of bounce rates variable
hist(ec$BounceRates)
#creating a histogram of exit rates variable
hist(ec$ExitRates)
#creating a histogram of page values variable
hist(ec$PageValues)
#creating a histogram of special day variable
hist(ec$SpecialDay)
```
### Observations from univariate analysis

* May has the most web page visits.

* Type 2 operating systenm is the most used.

* Browser type 2 is the most used.

* The highest number of individuals in the dataset came from region 1

* Traffic type 2 got most entries.

* Returning visitors formed the larger portion of the data.

* More web page visits were made during the weekday.

* There was lesser revenue.


#6. Bivariate & Multivariate Analysis
```{r}
#finding the covariance
cov <- cov(ec[,unlist(lapply(ec, is.numeric))])
cov
#Finding correlation
cor <- cor(ec[, unlist(lapply(ec, is.numeric))])
cor
```

```{r}
#selecting the true values from the revenue column
revenue <- ec[ec$Revenue == 'TRUE',]
head(revenue)
dim(revenue)
```

```{r}
#comparison between month and revenue brought in
#colnames(ec)
#frequency table of month
month1.freq <- table(revenue$Month)
sort(month1.freq, decreasing = TRUE)[1:5]
#Bar chart to show frequency distribution of months 
options(repr.plot.width = 10, repr.plot.height = 10)
barplot(c(month1.freq), main="A barchart of months.",
        xlab="months",
        ylab="revenue",
        cex.main=2, cex.lab=1.7,cex.sub=1.2,
        col=c("blue"))
#comparison between operating systems and revenue brought in
#frequency table of Operating systems
os1.freq <- table(revenue$OperatingSystems)
sort(os1.freq, decreasing = TRUE)[1:5]
#Bar chart to show frequency distribution of operating systems
options(repr.plot.width = 10, repr.plot.height = 10)
barplot(c(os1.freq), main="A barchart of operating systems.",
        xlab="operating systems",
        ylab="revenue",
        cex.main=2, cex.lab=1.7,cex.sub=1.2,
        col=c("green"))
#comparison between browser and revenue brought in
#frequency table of browser
browser1.freq <- table(revenue$Browser)
sort(browser1.freq, decreasing = TRUE)[1:5]
#Bar chart to show frequency distribution of browsers 
options(repr.plot.width = 10, repr.plot.height = 10)
barplot(c(browser1.freq), main="A barchart of browser.",
        xlab="browser",
        ylab="revenue",
        cex.main=2, cex.lab=1.7,cex.sub=1.2,
        col=c("blue"))
#comparison between region and revenue brought in
#frequency table of region
region1.freq <- table(revenue$Region)
sort(region1.freq, decreasing = TRUE)[1:5]
#Bar chart to show frequency distribution of regions 
options(repr.plot.width = 10, repr.plot.height = 10)
barplot(c(region1.freq), main="A barchart of regions.",
        xlab="region",
        ylab="revenue",
        cex.main=2, cex.lab=1.7,cex.sub=1.2,
        col=c("green"))
#comparison between traffic type and revenue brought in
#frequency table of traffic type
traffic1.freq <- table(revenue$TrafficType)
sort(traffic1.freq, decreasing = TRUE)[1:5]
#Bar chart to show frequency distribution of traffic type 
options(repr.plot.width = 10, repr.plot.height = 10)
barplot(c(traffic1.freq), main="A barchart of traffic type.",
        xlab="traffic type",
        ylab="revenue",
        cex.main=2, cex.lab=1.7,cex.sub=1.2,
        col=c("blue"))
#comparison between visitor type and revenue brought in
#frequency table of visitor type
visitor1.freq <- table(revenue$VisitorType)
sort(visitor1.freq, decreasing = TRUE)[1:5]
#Bar chart to show frequency distribution of visitor type 
options(repr.plot.width = 10, repr.plot.height = 10)
barplot(c(visitor1.freq), main="A barchart of visitor types.",
        xlab="visitor types",
        ylab="revenue",
        cex.main=2, cex.lab=1.7,cex.sub=1.2,
        col=c("green"))
#comparison between weekend and revenue brought in
#frequency table of weekend
weekend1.freq <- table(revenue$Weekend)
sort(weekend1.freq, decreasing = TRUE)[1:5]
#Bar chart to show frequency distribution of weekend 
options(repr.plot.width = 10, repr.plot.height = 10)
barplot(c(weekend1.freq), main="A barchart of weekend.",
        xlab="weekend",
        ylab="revenue",
        cex.main=2, cex.lab=1.7,cex.sub=1.2,
        col=c("blue"))
```

```{r}
#Pair plots
pairs(ec[, 1:10])
#Plotting a heat map using the correlation matrix
heatmap(x=cor, symm=TRUE)
#Plotting a correlogram
library('corrplot')
corrplot(cor, type='upper', order='hclust', tl.color='black', tl.srt=45)
```

### Observations from Bivariate and multivariate analysis

* Most revenue was made in the month of May. This would be because it also had the most web page visits.

* Operating system type 2 brought in more revenue.

* Users of browser type 2 also brought in more revenue.

* A great portion of revenue was gotten from region 1

* The traffic type that brought in most revenue was type 2

* Returning visitors also brought in more revenue as compared to the new and other visitors.

* A larger amount of revenue was made during the weekdays.

* Bounce rates and exit rates are highly correlated

* The number of product related websites visited is highly correlated to the product related duration spent.


#7. Implementing the Solution

## Feature engineering
```{r}
#removing the revenue column
ec.new <- ec[,1:17]
head(ec.new)
ec.rev <- ec$Revenue
ec.new[,12:15] <- sapply(ec.new[,12:15], as.character)
ec.new[,12:15] <- sapply(ec.new[,12:15], as.numeric)
head(ec.new)
```

```{r}
#one hot encoding factor columns
library(caret)
dmy = dummyVars(" ~ .", data = ec.new)
ec.encod = data.frame(predict(dmy, newdata = ec.new))
str(ec.encod)
dim(ec.encod)
```

```{r}
#normalizing data
normalize <- function(x){
  return ((x-min(x)) / (max(x)-min(x)))
}
normalize(ec.encod)
```

```{r}
# Applying the K-means clustering algorithm with no. of centroids(k)=3
# ---
# 
result<- kmeans(ec.encod,10) 
# Previewing the no. of records in each cluster
# 
result$size
result$cluster
result$centers
```
```{r}
# Verifying the results of clustering
# ---
# 
par(mfrow = c(2,2), mar = c(5,4,2,2))
# Plotting to see how various variable data points have been distributed in clusters
plot(ec.encod[c(1,2,3,4,5)], col = result$cluster)
```

## Hierachical clustering
```{r}
# First we use the dist() function to compute the Euclidean distance between observations, 
# d will be the first argument in the hclust() function dissimilarity matrix
# ---
#
d <- dist(ec.encod, method = "euclidean")
# We then hierarchical clustering using the Ward's method
# ---
# 
res.hc <- hclust(d, method = "ward.D2" )
res.hc
```

```{r}
# plotting the obtained dendrogram
# ---
# 
plot(res.hc, cex = 0.6, hang = -1)
```
#8. Challenging the solution

```{r}
#we'll use dbscan
library('dbscan')
# Applying our DBSCAN algorithm
# We want minimum 4 points with in a distance of eps(0.4)
# 
db<-dbscan(ec.encod,eps=0.4,MinPts = 4)
db
```

```{r}
# plotting our clusters as shown
# 
hullplot(ec.encod,db$cluster)
```

